Analytical Predictions of Explanet Obliquities Generated by Planet-Disk Interactions

### 2019-11-26 writeups/Su_disk_disp.tex
small notes changes

\title[Analytical Exoplanet Obliquities]{Analytical Predictions of Explanet
Obliquities Generated by Planet-Disk Interactions}
\author[Y. Su et\ al.]{
Yubo Su$^1$,
Dong Lai$^1$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
}

@@

Substantial planetary obliquities, defined as the angle between a planet's spin
and orbital axes, have featured in many astrophysical scenarios. They lie at the
heart of the obliquity tides scenario of radius enhancement in hot Jupiters
\citep{winn_otides, fabrycky_otides, millholland_wasp12b}. They have also been
invoked to explain peculiar thermal phase curves in hot Jupiters \citep[see
e.g.][]{millholland_signatures, ohno_infer_obl}. While exoplanet obliquities
have yet to be directly detected, techniques have been proposed to constrain
them \citep[see e.g.][]{schwarz_infer_obl, rauscher_infer_obl}.

@@

In the next two subsections, we further discuss theoretical background relevant
to our analysis of the evolution of this system. They are not essential to
reproducing the results presented in \autoref{ss:ad_ensemble}.

@@

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. The separatrix, three
zones, and their relations to the CSs are shown via a contour plot of the
Hamiltonian of the system
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}}.\label{eq:H}
\end{equation}
Such a contour plot is given in \autoref{fig:eq_1contours}. Note that
trajectories in zone $II$ librate about CS2 while those in zones $I$/$III$
circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ as given in
    \autoref{eq:H}, where warmer colors denote more positive values. The black
    solid line is the separatrix, which only exists for $\eta < \eta_{\rm c}$. The
    three zones, divided by the separatrix, are labeled. The Cassini states are
    labeled and have the same colors as \autoref{fig:cs_locs}. The interior of
    the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_{\rm c}$, but we may identify the phase space trajectories that flow into
    zone $II$ when evolved forward in time; this is the shaded region in the top
    left plot, bounded by the dotted black line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{subequations}\label{se:area_ward}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},\\
    \end{align}
    the areas for $\eta < \eta_{\rm c}$ are given by
    \begin{align}
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$, but a natural
extension exists: evolve an initial point $p$ under adiabatic decrease of $\eta
> \eta_{\rm c}$ until the separatrix appears at $\eta = \eta_{\rm c}$, then
identify $p$ with the zone it is in at $\eta_{\rm c}$. Since phase space area is
conserved under adiabatic evolution, this implies $A_i\p{\eta > \eta_{\rm c}} =
A_i(\eta_{\rm c})$. This extension is also reflected in the dashed boundary in
\autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by \autoref{se:area_ward}. Dotted lines correspond to small
    $\eta$ approximations discussed in \autoref{s:ad_approx}. Dashed lines for
    $\eta > \eta_{\rm c}$ effective values of $A_{\rm II}, A_{\rm III}$ for
    $\eta > \eta_{\rm c}$, denoting the area of points that would flow into
    either area under adiabatic decrease of $\eta$ from $\eta > \eta_{\rm
    c}$.}\label{fig:eq_areas}
\end{figure}

@@

A perturbation to a Hamiltonian system is considered \emph{adiabatic} if it is
sufficiently slow compared to all other relevant time scales in the problem.
In the current problem, the slowest time scale is the \emph{separatrix crossing
time scale}: as $\eta$ evolves, a trajectory may encounter the separatrix at
some value $\eta_\star$ which varies depending on its initial condition. If
$\eta$ changes sufficiently little during the separatrix-crossing orbit, the
evolution of the system may be considered adiabatic. The truly adiabatic limit
corresponds to $\epsilon \to 0$; we take $\epsilon = 3 \times 10^{-4}$ unless
otherwise noted.

One criterion for adiabaticity is that the separatrix crossing timescale be
longer than the the resonant libration period \citep{ward2004II}. The resonant
libration period for a nearly separatrix-crossing orbit is approximated by
the libration period near CS2 in \citet{millholland_disk}. We find that this
criterion performs remarkably well despire the expectation that the separatrix
crossing orbit should be much longer than libration near CS2 (approaching
infinity as $\epsilon \to 0$). One justification for this is that the libration
period about CS2 is the longest timescale in the system (see
\autoref{fig:lambda2})

This constraint can be computed
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{\rm cross} = \epsilon &\lesssim
            \frac{1}{T_{\rm lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta_\star\sin I \sin \theta_2
            \p{1 + \eta_\star \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{align}
where $\theta_2$ is the obliquity at CS2. This formula differs from that
presented in \citet{millholland_disk}, where the $\csc^3\theta_2$ term is
neglected and a square root is missing.

@@

\subsection{Transition to Non-adiabaticity}

@@

Later, these large scatters take on band-like structures that seem to persist
across values of $\theta_{\rm sd, i}$; we attribute these to the separatrix
crossing process becoming sensitive to the \emph{phase} of the final
libration/circulation orbit at separatrix crossing. This is equivalent to
separatrix crossing no longer being significantly slower than the resonant
libration period and violating the adiabatic assumption.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_25.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $\epsilon = 10^{-2.5}$,
    and restricting $\theta_{\rm sd, i} < 90^\circ$. A larger spread from the
    dynamical tracks is observed in the data.}\label{fig:3_ensemble_05_25}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_15.png}
    \caption{Same as \autoref{fig:3_ensemble_05_25} but for $\epsilon =
    10^{-1.5}$. Some small semblance of the evolutionary tracks remains, and the
    deviations appear to have a banded structure. This can be attributed to
    non-adiabaticity ``freezing-in'' the phase of the obliquity variations over
    the final libration/circulation orbit prior to separatrix
    crossing.}\label{fig:3_ensemble_05_15}
\end{figure}

A sample trajectory following in the style of \autoref{fig:ad_21} but for
$\epsilon = 0.3$ is provided in \autoref{fig:nonad_traj}. It is clear the
trajectory does not track level curves of the Hamiltonian during each individual
snapshot (e.g.\ the third snapshot).
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo_nonad.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo_nonad_subplots.png}
    \end{subfigure}
    \caption{Same as \autoref{fig:ad_21} but for a non-adiabatic $\epsilon =
    0.3$. In the top panel of the top plot, it is evident that the libration
    cycle about CS2 is unable to keep up with the swift motion of CS2 as $\eta$
    changes, decreasing the obliquity jump. In the bottom plot, we can see that
    individual orbits do not lie along level curves of the Hamiltonian, as the
    Hamiltonian phase space changes quickly compared to the period of
    circulation cycles.}\label{fig:nonad_traj}
\end{figure}

@@

A formula for $\theta_{\rm f}$ assuming $\theta_{\rm sd, i} = 0$ initially can
be given (see \autoref{s:ad_approx})
\begin{equation}
    \theta_{\rm f}\p{\theta_{\rm sd, i} = 0} = \sqrt{\frac{2\pi}{\epsilon}} \tan
        I\cos I.\label{eq:nonad_q_f}
\end{equation}
We can naively generalize this by recognizing that any nonzero $\theta_{\rm sd,
i}$ manifests as obliquity variations as $\hat{s}$ librates about $\hat{l}_{\rm
d}$. These can be accounted for by simply allowing the final obliquity to vary
over the same range as the initial obliquity, or,
\begin{equation}
    \theta_{\rm f}\p{\theta_{\rm sd, i}} - \sqrt{\frac{2\pi}{\epsilon}} \tan
        I\cos I\in \s{-\theta_{\rm sd, i}, \theta_{\rm sd,i}}
        .\label{eq:nonad_q_f_dist}
\end{equation}

We present the results of simulations for using $\epsilon = 0.3$ in
\autoref{fig:nonad_3_ensemeble}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_05.png}
    \caption{$\theta_{\rm  f}\p{\theta_{\rm sd, i}}$ at $\epsilon = 0.3$, firmly
    in the non-adiabatic regime. Note the clear double-valuedeness has
    disappeared, as have distinct dynamical tracks. The red dotted line presents
    the analytical prediction given by
    \autoref{eq:nonad_q_f_dist}.}\label{fig:nonad_3_ensemeble}
\end{figure}

@@

It is of interest to note the leading order behavior of $\theta$ in the small
$\theta_{\rm sd, i}$ limit given by \autoref{eq:qf_21_approx} and
\autoref{eq:qf_23_approx}. Therefore, if $\theta_{\rm sd, i}$ is generally
small, as would be expected for peacefully forming planets, resonance capture
induced by an adiabatically dissipating protoplanetary disk is expected to
significantly narrow the final $\theta$ spread compared to the initial
$\theta_{\rm sd, i}$ spread.

@@

\subsection{Stability}

@@

\subsection{Consequences for Adiabaticity}\label{ss:careful_ad}

In \autoref{s:ad}, we said the adiabaticity criterion is given by comparison of
the separatrix crossing timescale to the resonant libration period
\citep{ward2004II}, and we corrected the formula presented by
\citet{millholland_disk}. Two further comments can be made:
\begin{enumerate}
    \item The adiabaticity criterion compares how quickly the
        separatrix-crossing orbit is completed to how quickly the separatrix
        location shifts. However, it's not obvious that the libration period
        about CS2 captures the duration of the separatrix-crossing orbit, which
        as $\epsilon \to 0$ formally diverges. Thus, this criterion can only be
        taken as a necessary but not sufficient criterion to guarantee accurate
        predictions of $A$ post-encounter.

        We speculate a formally sufficient criterion may be able to be
        formulated as $\Delta t \lesssim \frac{\Delta \eta}{\dot{\eta}}$, where
        $\Delta t$ is the length of the separatrix-crossing orbit, and $\Delta
        \eta$ is the change in $\eta$ over the separatrix-crossing orbit. Note
        for all $\epsilon > 0$ that $\Delta t < \infty$. The exact form of such
        a criterion is inessential to our result.

    \item A different adiabaticity criterion can be defined for
        $III$-originating trajectories, as the separatrix-crossing orbit is much
        shorter (see e.g.\ \autoref{fig:ad_31}). However, the resultant
        criterion becomes extremely strict for the $III \to I$ track, as it goes
        through very small $\theta$. That this track is least adiabatic track
        can be seen by comparing \autoref{fig:ad_ensemble} and
        \autoref{fig:3_ensemble_05_25}, where the $III \to I$ track develops the
        largest spread about the analytical prediction.
\end{enumerate}

>>>>

\title[Analytical Exoplanet Obliquities]{Analytical Predictions of Explanet
Obliquities Generated by Planet-Disk Interactions}
\author[Y. Su and D. Lai]{
Yubo Su$^1$,
Dong Lai$^1$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
}

@@

Substantial planetary obliquities, defined as the angle between a planet's spin
and orbital axes, have been proposed in many exoplanetary systems. They lie at
the heart of the obliquity tides mechanism of radius enhancement in hot Jupiters
\citep{winn_otides, fabrycky_otides, millholland_wasp12b}. They have also been
invoked to explain peculiar thermal phase curves in hot Jupiters \citep[see
e.g.][]{millholland_signatures, ohno_infer_obl}. While exoplanet obliquities
have yet to be directly detected, techniques have been proposed to constrain
them \citep[see e.g.][]{schwarz_infer_obl, rauscher_infer_obl}.

@@

In the next two subsections, we further discuss theoretical background relevant
to our analysis of the evolution of this system. They are not essential to
reproducing the results of \autoref{s:ad} and \autoref{s:nonad}.

@@

When $\eta < \eta_{\rm c}$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. In
\autoref{fig:eq_1contours}, we show the separatrix, the three zones, and their
relations to the CSs via a contour plot of the Hamiltonian of the system
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_{\rm d}}.\label{eq:H}
\end{equation}
Note that trajectories in zone $II$ librate about CS2 while those in zones
$I$/$III$ circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ as given in
    \autoref{eq:H}, where warmer colors denote more positive values. The black
    solid line is the separatrix, which only exists for $\eta < \eta_{\rm c}$.
    The three zones, divided by the separatrix, are labeled. The Cassini states
    are labeled and have the same colors as \autoref{fig:cs_locs}. The interior
    of the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_{\rm c}$, but we may identify the points in phase space that flow into
    zone $II$ when evolved forward in time; this is the shaded region in the top
    left plot, bounded by the dotted black line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},\\
    \end{align}
\end{subequations}
the areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$, but a natural
extension exists: evolve an initial point $p$ under adiabatic decrease of $\eta$
until the separatrix appears at $\eta = \eta_{\rm c}$, then identify $p$ with
the zone it is in at $\eta_{\rm c}$. Since phase space area is conserved under
adiabatic evolution, this implies $A_i\p{\eta > \eta_{\rm c}} = A_i(\eta_{\rm
c})$. The boundary between these extended zones is denoted by the dashed black
line in the top left panel of \autoref{fig:eq_1contours}, where otherwise no
separatrix should exist since $\eta > \eta_{\rm c}$.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional areas of each of the zones $A_{\rm i}(\eta) /
    4\pi$ as given by \autoref{se:area_ward}. Dotted lines correspond to small
    $\eta$ approximations discussed in \autoref{s:ad_approx}. Dashed lines for
    $\eta > \eta_{\rm c}$ effective values of $A_{\rm II}, A_{\rm III}$ for
    $\eta > \eta_{\rm c}$, denoting the area of points that would flow into
    either area under adiabatic decrease of $\eta$ from $\eta > \eta_{\rm
    c}$.}\label{fig:eq_areas}
\end{figure}

@@

If a Hamiltonian system is parameterized by some parameter that is varied, its
evolution is considered \emph{adiabatic} if the parameter's variation is slow
compared to the Hamiltonian dynamics. In the current problem, the period of the
longest orbit of the constant-$\eta$ Hamiltonian system is infinite, namely
along the separatrix. This is an essential feature of the problem, as otherwise
trajectories would never be able to experience separatrix crossings. But this
seemingly implies adiabaticity is impossible to define.

Instead of requiring the variation of $\eta$ be slower than all timescales in
the Hamiltonian dynamics, we seek a weaker but nevertheless effective criterion.
In practice, trajectories only cross the separatrix once, while they spend many
orbits inside one of the three zones and far from the separatrix as $\eta$
evolves. Thus, a \emph{weak adiabaticity criterion} is that the variation of
$\eta$ be slower than all circulation/libration timescales in the Hamiltonian
system. If this criterion is satisfied, then a given trajectory will be
adiabatic for almost all of its evolution save one or two separatrix-crossing
orbits.

As is shown in \autoref{ss:eigens}, the libration frequency about CS2 is slower
than that about CS1 or CS3. As such, it is the slowest characteristic frequency
in the system. The weak adiabaticity criterion can be recast as requiring that
$\eta$ vary more slowly than this libration period; this is exactly the
criterion used in \citep{ward2004II,millholland_disk}. This constraint can be
computed
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}} = \epsilon &\lesssim
            \frac{1}{T_{\rm lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta\sin I \sin \theta_2
            \p{1 + \eta \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{align}
where $\theta_2$ is the obliquity at CS2. This formula differs somewhat from
that presented in \citet{millholland_disk}, where the $\csc^3\theta_2$ term is
neglected and a square root is missing.

For $I = 5^\circ$, $\eta \sim \eta_c$ we compute $\epsilon \lesssim 0.085$ is
the requirement for weak adiabaticity. Since our criterion is only a weak
constraint on adiabaticity, we use $\epsilon = 3 \times 10^{-4}$ in our
``adiabatic'' simulations. We explore the consequences of increasingly
nonadiabatic evolution in \autoref{ss:transition}.

@@

\subsection{Transition to Non-adiabaticity}\label{ss:transition}

@@

Later, these large scatters take on band-like structures that seem to persist
across values of $\theta_{\rm sd, i}$; we attribute these to the separatrix
crossing process becoming sensitive to the \emph{phase} of the final
libration/circulation orbit at separatrix crossing, which implies that $\eta$
changes substantially over the separatrix crossing orbit. This is clearly a
non-adiabatic effect but is permitted by the weak adiabaticity criterion,
which only compares $\eta$ evolution to dynamics far from the separatrix.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_25.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $\epsilon = 10^{-2.5}$,
    and restricting $\theta_{\rm sd, i} < 90^\circ$. A larger spread from the
    dynamical tracks is observed in the data.}\label{fig:3_ensemble_05_25}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_15.png}
    \caption{Same as \autoref{fig:3_ensemble_05_25} but for $\epsilon =
    10^{-1.5}$. Some small semblance of the evolutionary tracks remains, and the
    deviations appear to have a banded structure. This can be attributed to
    non-adiabaticity ``freezing-in'' the phase of the obliquity variations over
    the final libration/circulation orbit prior to separatrix
    crossing.}\label{fig:3_ensemble_05_15}
\end{figure}

A sample trajectory following in the style of \autoref{fig:ad_21} but for
$\epsilon = 0.3$ (violating even weak adiabaticity) is provided in
\autoref{fig:nonad_traj}. It is clear the trajectory does not track level curves
of the Hamiltonian during each individual snapshot (e.g.\ the third snapshot).
This manifests when CS2 migrates more quickly than the trajectory can librate
about CS2, precisely a violation of the weak adiabaticity criterion.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo_nonad.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo_nonad_subplots.png}
    \end{subfigure}
    \caption{Same as \autoref{fig:ad_21} but for a non-adiabatic $\epsilon =
    0.3$. In the top panel of the top plot, it is evident that the libration
    cycle about CS2 is unable to keep up with the swift motion of CS2 as $\eta$
    changes, decreasing the obliquity jump. In the bottom plot, we can see that
    individual orbits do not lie along level curves of the Hamiltonian, as the
    Hamiltonian phase space changes quickly compared to the period of
    circulation cycles.}\label{fig:nonad_traj}
\end{figure}

@@

We now consider the general problem of how a state evolves when $\eta$ changes
non-adiabatically. In \autoref{s:ad_approx}, we derive a formula for
$\theta_{\rm f}$ assuming no initial spin-disk misalignment:
\begin{equation}
    \theta_{\rm f}\p{\theta_{\rm sd, i} = 0} = \sqrt{\frac{2\pi}{\epsilon}} \tan
        I\cos I.\label{eq:nonad_q_f}
\end{equation}

An intuitive guess at generalizing this assumes that any nonzero $\theta_{\rm
sd, i}$ manifests as obliquity variations as $\hat{s}$ librates about
$\hat{l}_{\rm d}$. These can be accounted for by simply allowing the final
obliquity to vary over the same range as the initial obliquity, or,
\begin{equation}
    \theta_{\rm f}\p{\theta_{\rm sd, i}} - \sqrt{\frac{2\pi}{\epsilon}} \tan
        I\cos I\in \s{-\theta_{\rm sd, i}, \theta_{\rm sd,i}}
        .\label{eq:nonad_q_f_dist}
\end{equation}
In \autoref{fig:nonad_3_ensemeble}, we show that this proves a remarkably
accurate when taking $\epsilon = 0.3$.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_05.png}
    \caption{$\theta_{\rm  f}\p{\theta_{\rm sd, i}}$ at $\epsilon = 0.3$, firmly
    in the non-adiabatic regime. Note the clear double-valuedeness has
    disappeared, as have distinct dynamical tracks. The red dotted line presents
    the analytical prediction given by
    \autoref{eq:nonad_q_f_dist}.}\label{fig:nonad_3_ensemeble}
\end{figure}

@@

It is of interest to note the leading order behavior of $\theta$ in the small
$\theta_{\rm sd, i}$ limit given by \autoref{eq:qf_21_approx} and
\autoref{eq:qf_23_approx}. Therefore, if $\theta_{\rm sd, i}$ is generally
small, as might be expected for planets that do not experience strong collisions
or scattering, resonance capture induced by an adiabatically dissipating
protoplanetary disk is expected to significantly narrow the final $\theta$
spread compared to the initial $\theta_{\rm sd, i}$ spread.

@@

\subsection{Stability and Frequency of Local Oscillations}\label{ss:eigens}

@@
